---
title: "Analysis"
author: "Sarah I. Murphy"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE}
## load packages
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(lme4)
library(car)
library(emmeans)
library(broom)
```

# Stage 1a
```{r,message=FALSE}
#import data
s1_df <- read_csv("stage1_cleaned_data.csv")
s1_df$test <-  factor(s1_df$test,
                      c("SP.MSC", "SP.TSC", "HHR.TSC", "STSE"),
                      labels = c("SP-MSC", "SP-TSC", "HHR-TSC", "STSE"))
```

### Stage 1a - Table 1
```{r}
#summary for Table 1; means and sd across all powders
summary_byTestPlatingMedia <- s1_df %>%
  group_by(test,plating,media) %>%
  summarize(n=n(),
            mean=mean(log_mean_count_wNA,na.rm=TRUE),
            sd=sd(log_mean_count_wNA,na.rm=TRUE),
            median=median(log_mean_count_wNA,na.rm = TRUE),
            min=min(log_mean_count_wNA,na.rm=TRUE),
            max=max(log_mean_count_wNA,na.rm=TRUE))

summary_byTestPlatingMedia_SPRno <- s1_df %>%
  filter(SPR=="no")%>%
  group_by(test,plating,media) %>%
  summarize(n=n(),
            mean=mean(log_mean_count_wNA,na.rm=TRUE),
            sd=sd(log_mean_count_wNA,na.rm=TRUE),
            median=median(log_mean_count_wNA,na.rm = TRUE),
            min=min(log_mean_count_wNA,na.rm=TRUE),
            max=max(log_mean_count_wNA,na.rm=TRUE))
```

### Stage 1a - ANOVA
*Table 2
```{r}
m <- lm(log_mean_count~powder+plating*media*test+mean_spread_percent,data=s1_df)

#VIFs were investigated by fitting with lm() and using the car::vif function
vif(m) #vifs OK

#Additional model diagnostics were performed & looked OK
par(mfrow=c(2,2)) # Change the panel layout to 2 x 2
plot(m)

#run anova
anova(m)
```

### Stage 1a - Spreading and swarming


```{r}
m_spread <- lmer(mean_spread_percent_wNA ~ (1|powder) + test*media*plating, data=s1_df,na.action=na.exclude)

lsm_summ <- emmeans(m_spread,~test|media+plating)

df_lsm_summ <- as.data.frame(lsm_summ)
```

* Pairwise comparisons
```{r}
mod.emm.pairs <- pairs(lsm_summ)
df_mod.emm.pairs<- tidy(mod.emm.pairs)
```

* Figure 1
```{r}
ggplot(df_lsm_summ, aes(x = plating, y = emmean, color = media)) +
  facet_grid( ~ test) +
  geom_hline(yintercept = 25,
             col = "black",
             size = 1.5) +
  geom_point(stat = "identity", position = position_dodge(.75)) +
  geom_errorbar(stat = "identity",
                aes(ymin = lower.CL, ymax = upper.CL),
                size = 1,
                width = .75,
                position = position_dodge(.75),
                na.rm = TRUE) +
  geom_jitter(aes(x = plating, y = mean_spread_percent_wNA, color = media),
              data = s1_df,
              na.rm = TRUE,
              pch = 4,
              position = position_jitterdodge(jitter.width = 0.75, dodge.width = .75)) +
  labs(color = "Media", x = "Plating method", y = "Percent of plate covered by spreading and swarming colonies (%)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 
```

### Stage 1a - r^2 analyses
* Table 3
```{r}
s1_df2 <- s1_df %>% 
  filter( !(grepl("TSAwS",media) & plating=="Spread" & test=="SP.TSC"),
          !(grepl("PCMA",media) & plating=="Spread" & test=="SP.TSC"),
          !(grepl("BHI",media) & plating=="Spread" & test=="SP.TSC"),
          !(grepl("SMA",media) & plating=="Pour" & test=="SP.TSC"),
          !(grepl("PCMA",media) & plating=="Pour" & test=="SP.TSC") )
s1_df2$combo<- paste(s1_df2$media, s1_df2$plating, s1_df2$test, sep="_")

s1_df2$combo_abbrv <- make.names(abbreviate(s1_df2$combo))

#is this doing anything useful other than reshaping?
all_wide <- s1_df2 %>%
  group_by(powder,combo_abbrv) %>%
  summarise(mean=mean(log_mean_count)) %>%
  spread(key=combo_abbrv,value=mean) %>%
  ungroup() %>%
  data.frame() 
all_wide <- all_wide[c(2:44)]

SP.TSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="SP.TSC"])
HHR.TSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="HHR.TSC"])
STSE <- unique(s1_df2$combo_abbrv[s1_df2$test=="STSE"])
SP.MSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="SP.MSC"])

#measure test choices
abbrv_combo_lookup <- unique(s1_df2[,c("test","media","plating","combo_abbrv")])

#BHI
#test_combo <- c("BHI_P_H","BHI_P_SP.M","BHI_P_SP.T","BHI_P_ST")
#test_combo <- c("BHI_SO_H","BHI_SO_SP.M","BHI_SO_SP.T","BHI_SO_ST")
#test_combo <- c("BHI_S_H","BHI_S_SP","BHI_S_ST")

#PCMA
#test_combo <- c("PCMA_P_H","PCMA_P_SP","PCMA_P_ST")
#test_combo <- c("PCMA_SO_H","PCMA_SO_SP.M","PCMA_SO_SP.T","PCMA_SO_ST")
#test_combo <- c("PCMA_S_H","PCMA_S_SP","PCMA_S_ST")

#SMA
#test_combo <- c("PCA_P_H","PCA_P_SP","PCA_P_ST")
#test_combo <- c("PCA_SO_H","PCA_SO_SP.M","PCA_SO_SP.T","PCA_SO_ST")
#test_combo <- c("PCA_S_H","PCA_S_SP_M","PCA_S_SP_T","PCA_S_ST")

#TSAwS
test_combo <- c("TSAS_P_H","TSAS_P_SP.M","TSAS_P_SP.T","TSAS_P_ST") #check this
#test_combo <- c("TSAS_SO_H","TSAS_SO_SP.M","TSAS_SO_SP.T","TSAS_SO_ST")
#test_combo <- c("TSAS_S_H","TSAS_S_SP","TSAS_S_ST")

results <- data.frame(combo_abbrv=c(),Rsquared=c())
for( i in colnames(all_wide)) {
  if( !(i %in% test_combo) ) {
    form <- paste(i,"~",paste(test_combo,collapse="+"))
    m <- lm(form,data=all_wide)
    rsquared <- summary(m)$r.squared
    results <- rbind(results,data.frame(combo_abbrv=i,Rsquared=rsquared))
  }
}
# by test
results %>%
  left_join(abbrv_combo_lookup) %>%
  group_by(test) %>%
  summarize(mean.rsq=mean(Rsquared)) -> mean.rsq.byTest
mean.rsq.byTest

all_var <- s1_df2 %>%
  group_by(test,plating,media) %>%
  summarize(var=var(log_mean_count))

# percent variance
results %>%
  left_join(abbrv_combo_lookup) %>%
  left_join(all_var) %>%
  mutate(var_explained=Rsquared*var)  %>%
  summarize(prop_explained=sum(var_explained)/sum(var)) -> prop_explained
prop_explained

results %>%
  left_join(abbrv_combo_lookup) %>%
  left_join(all_var) %>%
  filter(test != "SP_MSC") %>%
  mutate(var_explained=Rsquared*var)  %>%
  summarize(prop_explained_no_msc=sum(var_explained)/sum(var)) -> prop_explained_withoutMSC
prop_explained_withoutMSC
```

```{r}
s1_df2 <- s1_df %>% 
  filter( !(grepl("TSAwS",media) & plating=="Spread" & test=="SP-TSC"),
          !(grepl("PCMA",media) & plating=="Spread" & test=="SP-TSC"),
          !(grepl("BHI",media) & plating=="Spread" & test=="SP-TSC"),
          !(grepl("SMA",media) & plating=="Pour" & test=="SP-TSC"),
          !(grepl("PCMA",media) & plating=="Pour" & test=="SP-TSC") )
s1_df2$combo<- paste(s1_df2$media, s1_df2$plating, s1_df2$test, sep="_")

s1_df2$combo_abbrv <- make.names(abbreviate(s1_df2$combo))

#is this doing anything useful other than reshaping?
all_wide <- s1_df2 %>%
  group_by(powder,combo_abbrv) %>%
  summarise(mean=mean(log_mean_count)) %>%
  spread(key=combo_abbrv,value=mean) %>%
  ungroup() %>%
  data.frame() 
all_wide <- all_wide[c(2:44)]

SP.TSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="SP.TSC"])
HHR.TSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="HHR.TSC"])
STSE <- unique(s1_df2$combo_abbrv[s1_df2$test=="STSE"])
SP.MSC <- unique(s1_df2$combo_abbrv[s1_df2$test=="SP.MSC"])

#measure test choices
abbrv_combo_lookup <- unique(s1_df2[,c("test","media","plating","combo_abbrv")])

#BHI
BHI_P <- c("BHI_P_H","BHI_P_SP.M","BHI_P_SP.T","BHI_P_ST")
BHI_SO <- c("BHI_SO_H","BHI_SO_SP.M","BHI_SO_SP.T","BHI_SO_ST")
BHI_S <- c("BHI_S_H","BHI_S_SP","BHI_S_ST")

#PCMA
PCMA_P <- c("PCMA_P_H","PCMA_P_SP","PCMA_P_ST")
PCMA_SO  <- c("PCMA_SO_H","PCMA_SO_SP.M","PCMA_SO_SP.T","PCMA_SO_ST")
PCMA_S <- c("PCMA_S_H","PCMA_S_SP","PCMA_S_ST")

#SMA
SMA_P <- c("SMA_P_H","SMA_P_SP","SMA_P_ST")
SMA_SO <- c("SMA_SO_H","SMA_SO_SP.M","SMA_SO_SP.T","SMA_SO_ST")
SMA_S <- c("SMA_S_H","SMA_S_SP.M","SMA_S_SP.T","SMA_S_ST")

#TSAwS
TSAS_P <- c("TSAS_P_H","TSAS_P_SP.M","TSAS_P_SP.T","TSAS_P_ST") #check this
TSAS_SO <- c("TSAS_SO_H","TSAS_SO_SP.M","TSAS_SO_SP.T","TSAS_SO_ST")
TSAS_S <- c("TSAS_S_H","TSAS_S_SP","TSAS_S_ST")

test_combo_vec <- c(BHI_P, BHI_SO, BHI_S,
                     PCMA_P, PCMA_SO, PCMA_S,
                     SMA_P, SMA_SO, SMA_S,
                     TSAS_P, TSAS_SO, TSAS_S)

results <- data.frame(combo_abbrv=c(),Rsquared=c(),testcombo=c())

for( j in test_combo_vec) {
for( i in colnames(all_wide)) {
  if( !(i %in% j) ) {
    form <- paste(i,"~",paste(j,collapse="+"))
    m <- lm(form,data=all_wide)
    rsquared <- summary(m)$r.squared
    results <- rbind(results,data.frame(combo_abbrv=i,Rsquared=rsquared,testcombo=j))
  }
}}


results %>%
  left_join(abbrv_combo_lookup) %>%
  group_by(testcombo,test) %>%
  summarize(mean.rsq=mean(Rsquared)) %>%
  spread(key=test,value=mean.rsq) %>%
  ungroup() %>%
  data.frame() -> mean.rsq.byTest
mean.rsq.byTest

# results %>%
#   left_join(abbrv_combo_lookup) %>%
#   group_by(test) %>%
#   summarize(mean.rsq=mean(Rsquared)) %>%
#   spread(key=test,value=mean.rsq) %>%
#   ungroup() %>%
#   data.frame() -> mean.rsq.byTest
# mean.rsq.byTest

all_var <- s1_df2 %>%
  group_by(test,plating,media) %>%
  summarize(var=var(log_mean_count))

```

```{r}
# percent variance

test_combo_vec <- c(PCMA_S) #run this indivudually for each of the test_combo

prop_df <- data.frame(testcombo=c(),prop_explain=c())
for (j in test_combo_vec){ 
results %>%
  filter(testcombo==j)%>%
  left_join(abbrv_combo_lookup) %>%
  left_join(all_var) %>%
  mutate(var_explained=Rsquared*var)  %>%
  summarize(prop_explained=sum(var_explained)/sum(var)) -> prop_explained
  prop_df <- rbind(prop_df, data.frame(testcombo=j,prop_explain=prop_explained))}

results %>%
  left_join(abbrv_combo_lookup) %>%
  left_join(all_var) %>%
  filter(test != "SP_MSC") %>%
  mutate(var_explained=Rsquared*var)  %>%
  summarize(prop_explained_no_msc=sum(var_explained)/sum(var)) -> prop_explained_withoutMSC
```

# Stage 1b
```{r,message=FALSE}
#import data
s1b_df <- read_csv("stage1b_data.csv")
```

### Stage 1b Bartlett test of homogeneity of variances
*Figure 2
```{r}
#SP_MSC
d_msc <- s1b_df %>% filter(test=="SP MSC")
# Center within powder/scheme
d_msc_centered <- d_msc %>%
  group_by(powder,sampling_scheme) %>%
  mutate(powder_mean=mean(log_cfu_g)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_cfu_g - powder_mean)
d_msc_centered_summary <- d_msc_centered %>%
  group_by(sampling_scheme) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_msc_centered_summary %>% ggplot(aes(x=sampling_scheme,y=mean)) + 
  labs(x="Sampling Scheme",y="Spore Count Deviance (log CFU/g)") +
  scale_color_discrete(guide=F) +
  geom_point(data=d_msc_centered,
             aes(x=sampling_scheme,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red")+
  ylim(-1,1) -> msc
bart_msc <- bartlett.test(log_cfu_g_centered~sampling_scheme,data=d_msc_centered)
bart_msc

#HHR
d_hhr <- s1b_df %>% filter(test=="HHR")
# Center within powder/scheme
d_hhr_centered <- d_hhr %>%
  group_by(powder,sampling_scheme) %>%
  mutate(powder_mean=mean(log_cfu_g)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_cfu_g - powder_mean)
d_hhr_centered_summary <- d_hhr_centered %>%
  group_by(sampling_scheme) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_hhr_centered_summary %>% ggplot(aes(x=sampling_scheme,y=mean)) + 
  labs(x="Sampling Scheme",y="Spore Count Deviance (log CFU/g)") +
  scale_color_discrete(guide=F) +
  geom_point(data=d_hhr_centered,
             aes(x=sampling_scheme,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red")+
  ylim(-1,1) -> hhr
bart_hhr <- bartlett.test(log_cfu_g_centered~sampling_scheme,data=d_hhr_centered)
bart_hhr

#STSE
d_stse <- s1b_df %>% filter(test=="STSE")
# Center within powder/scheme
d_stse_centered <- d_stse %>%
  group_by(powder,sampling_scheme) %>%
  mutate(powder_mean=mean(log_cfu_g)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_cfu_g - powder_mean)
d_stse_centered_summary <- d_stse_centered %>%
  group_by(sampling_scheme) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_stse_centered_summary %>% ggplot(aes(x=sampling_scheme,y=mean)) + 
  labs(x="Sampling Scheme",y="Spore Count Deviance (log CFU/g)") +
  scale_color_discrete(guide=F) + #getting rid of this gives us legend
  geom_point(data=d_stse_centered,
             aes(x=sampling_scheme,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red")+
  ylim(-1,1)-> stse
bart_stse <- bartlett.test(log_cfu_g_centered~sampling_scheme,data=d_stse_centered)
bart_stse

#arrange figures
ggarrange(msc,
          hhr,
          stse,
          labels=c("SP-MSC", "HHR-TSC", "STSE"),
          ncol = 3, nrow = 1)

##Make summary plots
ggplot(s1b_df, aes(x=powder,y=log_cfu_g))+
  geom_boxplot()+
  geom_jitter(data=s1b_df,
             aes(x=powder,y=log_cfu_g,color=sampling_scheme),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  facet_grid(test~sampling_scheme)
```

```{r}
#clear global environment
rm(list = ls())
```

# Stage 2

```{r,message=FALSE}
#import data
s2_df <- read_csv("stage2_data.csv")
s2_df2 <- read_csv("stage2_data_notImputed.csv")

#
df_PCMA <- s2_df %>%
  filter(media=="PCMA")
df_TSAwS <- s2_df %>%
  filter(media=="TSAwS")
```

### Stage 2 Bartlett test of homogeneity of variances
*Figure 3
```{r}
###PCMA analysis

#SP_MSC
d_PCMA_msc <- df_PCMA %>% filter(test=="SP.MSC")
# Center within powder/scheme
d_PCMA_msc_centered <- d_PCMA_msc %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_msc_centered_summary <- d_PCMA_msc_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_msc_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_msc_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_msc
bart_PCMA_msc <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_msc_centered) #we centered all log msc be respective powder
bart_PCMA_msc

#HHR_TSC
d_PCMA_hhr <- df_PCMA %>% filter(test=="HHR.TSC")
# Center within powder/scheme
d_PCMA_hhr_centered <- d_PCMA_hhr %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_hhr_centered_summary <- d_PCMA_hhr_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_hhr_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_hhr_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_hhr
bart_PCMA_hhr <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_hhr_centered)
bart_PCMA_hhr

#STSE
d_PCMA_stse <- df_PCMA %>% filter(test=="STSE")
# Center within powder/scheme
d_PCMA_stse_centered <- d_PCMA_stse %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_stse_centered_summary <- d_PCMA_stse_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_stse_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_stse_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_stse
bart_PCMA_stse <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_stse_centered)
bart_PCMA_stse
###TSAwS analysis

#SP_MSC
d_TSAwS_msc <- df_TSAwS %>% filter(test=="SP.MSC")
# Center within powder/scheme
d_TSAwS_msc_centered <- d_TSAwS_msc %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_msc_centered_summary <- d_TSAwS_msc_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_msc_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_msc_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_msc
bart_TSAwS_msc <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_msc_centered)
bart_TSAwS_msc


#HHR_TSC
d_TSAwS_hhr <- df_TSAwS %>% filter(test=="HHR.TSC")
# Center within powder/scheme
d_TSAwS_hhr_centered <- d_TSAwS_hhr %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_hhr_centered_summary <- d_TSAwS_hhr_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_hhr_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_hhr_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_hhr
bart_TSAwS_hhr <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_hhr_centered)
bart_TSAwS_hhr

#STSE
d_TSAwS_stse <- df_TSAwS %>% filter(test=="STSE")
# Center within powder/scheme
d_TSAwS_stse_centered <- d_TSAwS_stse %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_stse_centered_summary <- d_TSAwS_stse_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_stse_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-3,3)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_stse_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_stse
bart_TSAwS_stse <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_stse_centered)
bart_TSAwS_stse

#arrange figures
ggarrange(PCMA_msc, 
          TSAwS_msc,
          PCMA_hhr,
          TSAwS_hhr,
          PCMA_stse,
          TSAwS_stse,
          labels=c("PCMA SP-MSC", "TSAwS SP-MSC", "PCMA HHR-TSC", "TSAwS HHR-TSC", "PCMA STSE", "TSAwS STSE"),
          ncol = 2, nrow = 3)
```

### Stage 2 Summary
* Supplemental Figure S2 [w/out imputed values]
* Means across labs by test for PCMA & TSAwS
```{r}
##Make summary plots
#plot without imputed values
s2_df2$mean_count2 <- s2_df2$mean_count
s2_df2$mean_count2[s2_df2$mean_count2==0]<-10
s2_df2$log_mean_count <- log10(s2_df2$mean_count2)
# 
s2_df2$lab <- factor(s2_df2$lab,c("UA","UB","UC","UD"))
s2_df2$test <-  factor(s2_df2$test,
                      c("SP.MSC", "HHR.TSC", "STSE"),
                      labels = c("SP-MSC", "HHR-TSC", "STSE"))

ggplot(s2_df2, aes(x=powder,y=log_mean_count))+
  geom_boxplot()+
  geom_point(data=s2_df2,
             aes(x=powder,y=log_mean_count,color=lab),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  ylim(0,5)+
  labs(x="Powder",y="Spore Count (log CFU/g)") +
  facet_grid(test~media)

#Get means across labs without imputed values
summary_s2_df2b <- s2_df2 %>%
  group_by(media,test,lab)%>%
  summarize(mean=mean(log_mean_count),
            sd=sd(log_mean_count))
```

```{r}
#clear global environment
rm(list = ls())
```

# Stage 3
```{r,message=FALSE}
#import data
s3_df <-read_csv("stage3_data.csv")

#prep data
df_PCMA <- s3_df %>%
  filter(media=="PCMA")
df_TSAwS <- s3_df %>%
  filter(media=="TSA")
```

### Stage 3 Bartlett test of homogeneity of variances
* Figure 4
```{r}
###PCMA analysis
#SP_MSC
d_PCMA_msc <- df_PCMA %>% filter(test=="SP.MSC")
# Center within powder/scheme
d_PCMA_msc_centered <- d_PCMA_msc %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_msc_centered_summary <- d_PCMA_msc_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_msc_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_msc_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_msc
bart_PCMA_msc <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_msc_centered)
bart_PCMA_msc


#HHR_TSC
d_PCMA_hhr <- df_PCMA %>% filter(test=="HHR.TSC")
# Center within powder/scheme
d_PCMA_hhr_centered <- d_PCMA_hhr %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_hhr_centered_summary <- d_PCMA_hhr_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_hhr_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_hhr_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_hhr
bart_PCMA_hhr <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_hhr_centered)
bart_PCMA_hhr

#STSE
d_PCMA_stse <- df_PCMA %>% filter(test=="STSE")
# Center within powder/scheme
d_PCMA_stse_centered <- d_PCMA_stse %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_PCMA_stse_centered_summary <- d_PCMA_stse_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_PCMA_stse_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_PCMA_stse_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> PCMA_stse
bart_PCMA_stse <- bartlett.test(log_cfu_g_centered~lab,data=d_PCMA_stse_centered)
bart_PCMA_stse 

###
###TSAwS analysis

#SP_MSC
d_TSAwS_msc <- df_TSAwS %>% filter(test=="SP.MSC")
# Center within powder/scheme
d_TSAwS_msc_centered <- d_TSAwS_msc %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_msc_centered_summary <- d_TSAwS_msc_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_msc_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_msc_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_msc
bart_TSAwS_msc <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_msc_centered)
bart_TSAwS_msc


#HHR_TSC
d_TSAwS_hhr <- df_TSAwS %>% filter(test=="HHR.TSC")
# Center within powder/scheme
d_TSAwS_hhr_centered <- d_TSAwS_hhr %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_hhr_centered_summary <- d_TSAwS_hhr_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_hhr_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_hhr_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_hhr
bart_TSAwS_hhr <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_hhr_centered)
bart_TSAwS_hhr 

#STSE
d_TSAwS_stse <- df_TSAwS %>% filter(test=="STSE")
# Center within powder/scheme
d_TSAwS_stse_centered <- d_TSAwS_stse %>%
  group_by(powder) %>%
  mutate(powder_mean=mean(log_mean_count)) %>%
  ungroup() %>%
  mutate(log_cfu_g_centered=log_mean_count - powder_mean)
d_TSAwS_stse_centered_summary <- d_TSAwS_stse_centered %>%
  group_by(lab) %>%
  summarize(mean=mean(log_cfu_g_centered),sd=sd(log_cfu_g_centered))
d_TSAwS_stse_centered_summary %>% ggplot(aes(x=lab,y=mean)) + 
  labs(x="Lab",y="Spore Count Deviance (log CFU/g)") +
  ylim(-2,2)+
  scale_color_discrete(guide=F) +
  geom_point(data=d_TSAwS_stse_centered,
             aes(x=lab,y=log_cfu_g_centered,color=powder),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
  #theme(text=element_text(family="Times New Roman", size=12)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean - sd,
                    ymax=mean + sd),
                width=0.3,
                color="red") -> TSAwS_stse
bart_TSAwS_stse <- bartlett.test(log_cfu_g_centered~lab,data=d_TSAwS_stse_centered)
bart_TSAwS_stse

ggarrange(PCMA_msc, 
          TSAwS_msc,
          PCMA_hhr,
          TSAwS_hhr,
          PCMA_stse,
          TSAwS_stse,
          labels=c("PCMA SP-MSC", "TSAwS SP-MSC", "PCMA HHR-TSC", "TSAwS HHR-TSC", "PCMA STSE", "TSAwS STSE"),
          ncol = 2, nrow = 3)
```

### Stage 3 Summary
* Supplemental Figure S3
* Means across labs by test for PCMA & TSAwS
```{r}
##Make summary plots
s3_df$test <-  factor(s3_df$test,
                      c("SP.MSC", "HHR.TSC", "STSE"),
                      labels = c("SP-MSC", "HHR-TSC", "STSE"))

ggplot(s3_df, aes(x=powder,y=log_mean_count))+
  geom_boxplot()+
  ylim(0,4)+
  geom_point(data=s3_df,
             aes(x=powder,y=log_mean_count,color=lab),
             position=position_jitterdodge(jitter.width = 0.05,dodge.width=0.7)) +
    labs(x="Powder",y="Spore Count (log CFU/g)") +
  facet_grid(test~media)

#Get means across labs
summary_s3_dfb <- s3_df %>%
  group_by(media,test,lab)%>%
  summarize(mean=mean(log_mean_count),
            sd=sd(log_mean_count))
```

```{r}
#clear global environment
rm(list = ls())
```
